The symmetry problem of the \textit{Fermi} and eROSITA bubbles: A proof-of-concept study

### 2022-02-13 paper.tex
update

Early attempts \citep{Sarkar2015,Yang2017,Zhang2020}\
to model the symmetric \textit{Fermi} and eROSITA bubbles (collectively we call the Galactic bubbles)\
assumed that the jet is vertical to the Galactic plane. However,\
observation \citep{Gallimore2006} has proposed that\
there is a lack of preferred orientation of jets with respect to a disc normal.\
In contrast, a number of galaxies in which the jets are tilted to the disc normal\
(e.g. NGC 3079, \citealt{Cecil2001}; NGC 1052, \citealt{Dopita2015}),\
including galaxies in which the jets lie in the plane of the disc (e.g. IC 5063, \citealt{Morganti2015}).

To this end, this work introduces a dense thin disc of interstellar medium\
around a jet source to deflects and decelerates the inclined jets,\
in an attempt to remedy the symmetry problem of the Galactic bubbles.\

@@

@@

 \Cref{fig__jetI5+ismSeed3-45deg_no-disc-stratified-ambient-Pres-Dens-Temp} shows\
 the slices of pressure (top), temperature (middile), number density (bottom)\
 at the end of simulation $t=12.39$ Myr.\
 The slices pass through the bipolar jet source injecting along $z=-y$ direction\
 for a duration $t=0$--$1.2$ Myr.\
 Comparison between the clumpy (first column from left)\
 and the smooth disc (second) in a stratified atmosphere\
 shows that the initial density distribution of the dense disc has an insignificant effect\
 on the overall dynamics of bubbles. However, the outermost bubbles arising from the smooth disc\
 in an uniform atmosphere (third) is spherical-shape,\
 suggesting the stratification facilitates the outermost bubbles elongation significantly.\
 Also, the two rightmost columns reveal that the development of the innermost bubbles (dashed box)\
 is always associated with the disc, and without the disc,\
 the outermost bubbles and the turbulent region will be tilted,
 indicating the dense disc is crucial for\
 the symmetry of the Galactic bubbles and for the innermost bubbles formation.

@@

  \begin{figure}
    \includegraphics[width=\columnwidth]{figures/fig__jetI5+ismSeed3-45deg_no-disc-stratified-ambient-Pres-Dens-Temp.png}
    \caption{
             The slices of pressure (top), temperature (middle), and number density (bottom)\
             at the end of simulation $t=12.39$ Myr.\
             The slices pass through a bipolar jet source injecting along $z=-y$ direction\
             for a duration $t=0$--$1.2$ Myr.\
             Comparison between the clumpy (first column from left)\
             and the smooth disc (second) in a stratified atmosphere\
             shows that the initial density distribution of the dense disc has an insignificant effect\
             on the overall dynamics of bubbles. However, the outermost bubbles arising from the smooth disc\
             in an uniform atmosphere (third) is spherical-shape,\
             suggesting the stratification facilitates the outermost bubbles elongation significantly.\
             Also, the two rightmost columns reveal that the development of the innermost bubbles (dashed box)\
             is always associated with the disc, and without the disc,\
             the outermost bubbles and the turbulent region will be tilted,
             indicating the dense disc is crucial for\
             the symmetry of the Galactic bubbles and for the innermost bubbles formation.
             }
    \label{fig__jetI5+ismSeed3-45deg_no-disc-stratified-ambient-Pres-Dens-Temp}
  \end{figure}

@@

  First, the broad agreement between simulated and observed X-ray map hint that\
  the full vertical extent of the eROSITA bubbles can be properly formed by an inclined jet within a thin disk\
  of dense interstellar medium.\

  Second, as shown in \Cref{fig__jetI5+ismSeed3-45deg_no-disc-stratified-ambient-Pres-Dens-Temp},\
  the half-width of the outermost bubbles is around 7 kpc,\
  corresponding to an half angular width $\sin^{-1}(7 \text{ kpc}/R_{\odot})\sim122^{\circ}$,\
  which is as wide as the eROSITA bubbles in the simulated X-ray map (top panel in \Cref{fig__xray_0.8keV_angle_000}).\
  We therefore suggest that the nature of eROSITA bubbles is indeed a forward shock\
  propagating through the halo gas, as previously proposed by \citet{Predehl2020}.

% 2. North Polar Spur is not shown as it is a supernova remanent near us.
  Third, the innermost bubbles, even though with high column density, is invisible in the simulated X-ray map as\
  the temperature of innermost bubbles is around $1$ -- $10$ eV\
  (see the temperature slice in \Cref{fig__jetI5+ismSeed3-45deg_no-disc-stratified-ambient-Pres-Dens-Temp}).\
  Consequently, the X-ray emission within the innermost bubbles
  is severely suppressed by the cutoff $\exp\left[-h\nu/k_{B}T\right]$ in the thermal Bremsstrahlung emissivity.\
  This is the reason why the innermost bubbles is unseen in the observation.

@@

\section{Conclusions}
\label{Conclusions}
We have found qualitative similarities between the results our simulations\
and the observed spectrum and morphological structure of the Galactic bubbles.

>>>>

Early attempts \citep{Sarkar2015,Yang2017,Zhang2020}\
to model the symmetric \textit{Fermi} and eROSITA bubbles (collectively we call the Galactic bubbles)\
assumed that the jet is vertical to the Galactic plane. However,\
observation \citep{Gallimore2006} has proposed that\
there is a lack of preferred orientation of jets with respect to\
the rotation axis of a galatic plane (disc normal, hereafter).\
In contrast, a number of galaxies in which the jets are tilted to the disc normal\
(e.g. NGC 3079, \citealt{Cecil2001}; NGC 1052, \citealt{Dopita2015}),\
including galaxies in which the jets lie in the plane of the disc (e.g. IC 5063, \citealt{Morganti2015}).

To this end, this work introduces a dense thin disc of interstellar medium\
around a jet source to deflects and decelerates the inclined jets,\
in an attempt to resolve the symmetry problem of the Galactic bubbles.\

@@

  %, and prevent the disk from destruction by .
  % ablate

@@

 \Cref{fig__jetI5+ismSeed3-45deg} shows\
 the slices of pressure (top), temperature (middile), number density (bottom)\
 at the end of simulation $t=12.39$ Myr.\
 The slices pass through the bipolar jet source injecting along $z=-y$ direction\
 for a duration $t=0$--$1.2$ Myr.\
 Comparison between the clumpy (first column from left)\
 and the smooth disc (second) in a stratified atmosphere\
 shows that the initial density distribution of the dense disc has an insignificant effect\
 on the overall dynamics of bubbles. However, the outermost bubbles arising from the smooth disc\
 in an uniform atmosphere (third) is spherical-shape,\
 suggesting the stratification facilitates the outermost bubbles elongation significantly.\
 Also, the two rightmost columns reveal that the development of the innermost bubbles (dashed box)\
 is always associated with the disc, and without the disc,\
 the outermost bubbles and the turbulent region will be tilted,
 indicating the dense disc is crucial for\
 the symmetry of the Galactic bubbles and for the innermost bubbles formation.

@@

  \begin{figure}
    \includegraphics[width=\columnwidth]{figures/fig__jetI5+ismSeed3-45deg.png}
    \caption{
             The slices of pressure (top), temperature (middle), and number density (bottom)\
             at the end of simulation $t=12.39$ Myr.\
             The slices pass through a bipolar jet source injecting along $z=-y$ direction\
             for a duration $t=0$--$1.2$ Myr.\
             Comparison between the clumpy (first column from left)\
             and the smooth disc (second) in a stratified atmosphere\
             shows that the initial density distribution of the dense disc has an insignificant effect\
             on the overall dynamics of bubbles. However, the outermost bubbles arising from the smooth disc\
             in an uniform atmosphere (third) is spherical-shape,\
             suggesting the stratification facilitates the outermost bubbles elongation significantly.\
             Also, the two rightmost columns reveal that the development of the innermost bubbles (dashed box)\
             is always associated with the disc, and without the disc,\
             the outermost bubbles and the turbulent region will be tilted,
             indicating the dense disc is crucial for\
             the symmetry of the Galactic bubbles and for the innermost bubbles formation.
             }
    \label{fig__jetI5+ismSeed3-45deg}
  \end{figure}

@@

  First, the broad agreement between simulated and observed X-ray maps hints that\
  the full vertical extent of the eROSITA bubbles can be properly formed by an inclined jet within a thin disk\
  of dense interstellar medium.\

  Second, as shown in \Cref{fig__jetI5+ismSeed3-45deg},\
  the half-width of the outermost bubbles is around 7 kpc,\
  corresponding to an half angular width $\sin^{-1}(7 \text{ kpc}/R_{\odot})\sim122^{\circ}$,\
  which is as wide as the eROSITA bubbles in the simulated X-ray map (top panel in \Cref{fig__xray_0.8keV_angle_000}).\
  We therefore suggest that the nature of eROSITA bubbles is indeed a forward shocks\
  that have been driven into the northern and the southern Galactic halo,
  as previously proposed by \citet{Predehl2020}.

% 2. North Polar Spur is not shown as it is a supernova remanent near us.
  Third, the innermost bubbles\
  shown in \Cref{fig__jetI5+ismSeed3-45deg},\
  even though with high column density, is invisible in the simulated X-ray map as\
  the temperature of innermost bubbles is around $1$ -- $10$ eV\
  (see the temperature slice in \Cref{fig__jetI5+ismSeed3-45deg}).\
  Consequently, the X-ray emission within the innermost bubbles
  is severely suppressed by the cutoff $\exp\left[-h\nu/k_{B}T\right]$ in the thermal Bremsstrahlung emissivity.\
  This is the reason why the innermost bubbles is unseen in the observation.

@@

\section{Conclusions}
\label{Conclusions}
The symmetry of the Galactic bubbles is worth exploring since observations\
show that there is no preference for the orientation of AGN jets relative to\
the rotation axis of a galactic plane.

In this work, we introduce a thin, dense disk composed of interstellar medium\
to divert the inclined jet at an angle $45^{\circ}$\
to the disk normal in the past energetic event of the central SMBH.

The jet

Also, We investigate the properties of the Galactic bubbles and the microwave\
haze using 3D special relativistic hydrodynamic simulations of CR injection from\
the SMBH assuming the leptonic model.

Our findings are as follows.

\begin{itemize}

\item The Galactic bubbles are nearly symmetric about the Galactic plane even the jet is at an angle $45^{\circ}$\
      with respect to the disc normal.
\item The randomness of clumpy disk has a insignificant impact on the overall dynamic of the Galactic bubbles.\
      The ambient stratification facilitates the eROSITA bubbles elongation significantly.
      The development of the inverse shocks are always associated with the dense disc.
\item The edge of the eROSITA bubbles is a forward shock,\
      originally driven by an inclined bipolar jets that emanated from the GC 12.39 million years ago,\
      and stretched by stratified ambient subsequently.
\item The nature of the \textit{Fermi} bubbles are essentially a turbulent flows\
      that can stochastically accelerate CRe. Associated with
      contact discontinuity\
\item

\end{itemize}

Over the past few years, there have been numerous efforts,\
both observational and theoretical, to uncover the nature of the Galactic bubbles.

Spektr-RG \citep{Sunyaev2021} and eROSITA \citep{Predehl2021}

We have found qualitative similarities between the results our simulations\
and the observed spectrum and morphological structure of the Galactic bubbles.
